Embedding Dimension of Polyhedral Losses

### 2020-01-30 colt-20/embed-dim.tex
my crack at most of a geometric proof

\begin{lemma} \label{lemma:abstain-v-h}
  For the abstain loss $\ell_{1/2}$, the level set of abstain satisfies $\gamma_{\bot} = \conv\{(\delta_y + \delta_{y'})/2 : y,y' \in \Y, y\neq y'\} = \{p : Bp \geq \vec{0}\}$ where $\delta_y$ puts probability one on $y$ and $B = \ones \ones^{\intercal} - 2I \in \reals^{5 \times 5}$, i.e. has entries $-1$ on the diagonal and $1$ everywhere else.
\end{lemma}
\begin{proof}
  Recall that $\gamma_{\bot}$ is the set of distributions $p$ with $\max_y p(y) = 1/2$.
  First, note that each distribution of the form $(1/2, 1/2, 0, 0, 0)$ and so on is in $\gamma_{\bot}$.
  Meanwhile, every such $p$ can be written as a convex combination of these corners.
  Second, note that if $p \in \gamma_{\bot}$, then $p(y) \leq 1/2$ for all $y \in \Y$.
  These constraints can be rewritten as $\inprod{p}{b} \geq 0$ where $b_y = -1$ and $b_{y'} = 1$ for all $y' \neq y$, literally requiring $p(y) \leq \sum_{y' \neq y} p(y')$.\bo{Check notation $p_y$ versus $p(y)$.}
\end{proof}

>>>>

\begin{lemma} \label{lemma:abstain-v-h}
  For the abstain loss $\ell_{1/2}$, the level set of abstain satisfies $\gamma_{\bot} = \conv\{(\delta_y + \delta_{y'})/2 : y,y' \in \Y, y < y'\} = \{p : Bp \geq \vec{0}\}$ where $\delta_y$ puts probability one on $y$ and $B = \ones \ones^{\intercal} - 2I \in \reals^{5 \times 5}$, i.e. has entries $-1$ on the diagonal and $1$ everywhere else.
\end{lemma}
\begin{proof}
  Recall that $\gamma_{\bot}$ is the set of distributions $p$ with $\max_y p(y) = 1/2$.
  First, note that each distribution of the form $(1/2, 1/2, 0, 0, 0)$ and so on is in $\gamma_{\bot}$.
  Meanwhile, every such $p$ can be written as a convex combination of these corners.
  Second, note that if $p \in \gamma_{\bot}$, then $p(y) \leq 1/2$ for all $y \in \Y$.
  These constraints can be rewritten as $\inprod{p}{b} \geq 0$ where $b_y = -1$ and $b_{y'} = 1$ for all $y' \neq y$, literally requiring $p(y) \leq \sum_{y' \neq y} p(y')$.\bo{Check notation $p_y$ versus $p(y)$.}
\end{proof}

\begin{observation} \label{obs:qfp-invariant-lin}
  For any invertible $A \in \reals^{d \times d}$, if $V,\{X^j : j \in [\ell]\}$ is a feasible solution to the quadratic feasibility program, then so is $(VA),\{A^{-1}X^j : j \in [\ell]\}$.
\end{observation}
\begin{proof}
  The halfspace constraints are $(VA)(A^{-1}X^j) \leq B \iff VX^{j} \leq B$.
  The $j$th vertex constraint is a vector equality $\sum_{y \in \Y} p^j(y) (A^{-1}X^j)_y = \vec{0}$.
  If we let $a_m$ be the $m$th row of $A^{-1}$, then the $m$th row of the vector equality is
  \begin{align*}
    0 &= \sum_{y \in \Y} p^j(y) \inprod{a_m}{x^j_y}  \\
      &= \inprod{a_m}{\sum_{y \in \Y} p^j(y) x^j_y}  \\
      &= 0
  \end{align*}
  so the program is feasible.
\end{proof}

\bo{I think around here we want a setup of the QFP where we introduce our specific notation like $x^{ij}_i$.}

\bo{We need to prove the dumb case that it won't be feasible to set $v_i = v_j$.}

\begin{corollary} \label{cor:qfp-wlog-v1}
  If there is a feasible solution to the quadratic feasibility program, then there is a feasible solution where $v_1$ is the first standard basis vector and $\|v_i\| \leq \|v_1\| = 1$ for all $i$.
\end{corollary}
\begin{proof}
  In particular, we can take a series of matrices $A$ in Observation \ref{obs:qfp-invariant-lin} that permutate the rows of $V$, scale\footnote{Note one can show $V = 0$ is not feasible unless $B$ is a trivial property, i.e. essentially has no rows at all.} all rows by $\frac{1}{\|v_1\|}$, and linearly map $v_1$ to $(1,0,\dots,0)$.
\end{proof}

\begin{lemma} \label{lemma:xiji-minus-xijj}
  In any feasible solution to the QFP for $\gamma_{\bot}$ and $\ell_{1/2}$, we have $x^{ij}_i = -x^{ij}_j$ for all $i<j$ in $\Y$.
\end{lemma}
\begin{proof}
  Directly from the vertex constraints: $p^{ij} = \frac{1}{2}\delta_i + \frac{1}{2}\delta_j$, so the $ij$ constraint reduces to $\frac{1}{2}x^{ij}_i + \frac{1}{2}x^{ij}_j = 0$.
\end{proof}

\begin{lemma} \label{lemma:unique-spot-xiji}
  For $d=2$, any pair of linearly independent $v_i,v_j$ rule out all except for a unique set of points $x^{ij}_i$ and $x^{ij}_j$ from being feasible.
\end{lemma}
\begin{proof}
  From the halfspace constraints, we must have $\inprod{v_i}{x^{ij}_i} \leq -1$ and $\inprod{v_i}{x^{ij}_j} \leq 1$, which combines with Lemma \ref{lemma:xiji-minus-xijj} to give $\inprod{v_i}{x^{ij}_i} = 1$.
  This immediately also gives $\inprod{v_j}{x^{ij}_i} = -1$.
  This system of two inequalities in two dimensions has exactly one solution if $v_i,v_j$ are linearly independent.
\end{proof}

\begin{lemma} \label{lemma:180-degree-no-three}
  There is no feasible solution to the QFP for $\gamma_{\bot}$ and $\ell_{1/2}$ where three vectors $v_i,v_j,v_m$ lie strictly within a halfspace through the origin (i.e. all within $180^{\circ}$ of each other).
\end{lemma}
\begin{proof}
  Let three of the vectors be given, lying strictly inside a halfspace, and label them clockwise clockwise as $v_1,v_2,v_3$.
  WLOG suppose $v_1$ points vertically ``up'', as in Figure \ref{bo:todo}.
  By Lemma \ref{lemma:unique-spot-xiji}, the possible locations of the following points are all uniquely determined: $x^{ij}_i, x^{ij}_j$ for $(i,j) \in \{(1,2),(1,3),(2,3)\}$.
  Both points $x^{12}_1$ and $x^{13}_1$ lie on the line $\inprod{v_1}{x} = -1$, i.e. a horizontal line below the origin.
  We have constraints $\inprod{v_2}{x^{12}_1} = 1$ and $\inprod{v_2}{x^{13}_1} \leq 1$.
  This implies $x^{13}_1$ is left of $x^{12}_1$ on the line.
  But the symmetric constraints $\inprod{v_3}{x^{13}_1} \leq 1$ and $\inprod{v_3}{x^{12}_1} \leq 1$ imply that $x^{12}_1$ is left of $x^{13}_1$ on the line.
  This implies we must have $x^{12}_1 = x^{13}_1$.

  If we consider the four lines $\inprod{v_2}{x}=1, \inprod{v_2}{x}=-1, \inprod{v_3}{x}=1, \inprod{v_3}{x}=-1$, we therefore have three points of intersection with the line $\inprod{v_1}{x}=1$ and three with the line $\inprod{v_1}{x}=-1$.
  WLOG, these points from top left to top right are: $x^{12}_2$ (which equals $x^{13}_3$), the intersection with $\inprod{v_2}{x}=1$, and the intersection with $\inprod{v_3}{x}=1$; and therefore from bottom left to bottom right are: the intersection with $\inprod{v_3}{x}=-1$, the intersection with $\inprod{v_2}{x}=-1$, and the intersection with $x^{12}_1$ (which equals $x^{13}_1$).
  \bo{Needs more justification why these points are always in this order (up to renaming of $v_2$ and $v_3$).}

  This implies that the lines $\inprod{v_2}{x}=-1$ and $\inprod{v_3}{x}=1$, in particular, do not intersect anywhere within the bounds of $\inprod{v_1}{x} \in [-1,1]$.
  Therefore, either their intersection point $x^{23}_2$ or its negative $x^{23}_3$ violates the feasibility constraint $\inprod{v_1}{x} \leq 1$.
  This proves there is no feasible solution.
\end{proof}

\begin{proposition}
	The abstain loss $\ell_{1/2}$ with $n=5$ is not $2$-embeddable.
\end{proposition}
\begin{proof}
  Let any $5$ vectors be given, numbered clockwise.
  $v_1,v_2,v_3$ cannot lie in a cone of strictly less than $180^{\circ}$, as this would contradict \bo{the previous}.
  So the clockwise angle between $v_1$ and $v_3$ is at least $180^{\circ}$.
  Since there are no duplicates, this implies that the clockwise angle between $v_4$ and $v_1$, which includes $v_5$, is strictly less than $180^{\circ}$.
  This contradicts \bo{the previous}.
\end{proof}
